Paper

### 2020-02-14 Paper.Rmd
updating for paper

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)
library(ggrepel)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(RColorBrewer)

@@

cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()
```

@@

load(file, verbose=F)
#dim(dysplasia.df)
#length(labels)

@@

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

@@

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

@@

error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')

@@

error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
summary(error$Abs.Error)
sd(error$Abs.Error)

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')

@@

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

@@

getPalette = colorRampPalette(brewer.pal(5, "Blues"))
p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
  geom_line(linetype='solid', size=0.5) + 
  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
  scale_x_reverse() +
  scale_color_manual(values=getPalette(length(unique(df$is)))) +
  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
p

@@

# Not sure why I did this
df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

ggplot() + 
  geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

@@

endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
  bind_rows(
    pROC::coords(rocEndoM, 'best'),
    pROC::coords(rocEndoA, 'best'),
    pROC::coords(rocLastEndo, 'best'),
    pROC::coords(rocFirstEndo, 'best'),
    pROC::coords(rocRandEndo, 'best')
  ))
```

@@

```{r}
pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% 
                                     map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))

@@

rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')

ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6, eval=F}
all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
  dplyr::rename('n.samples'='n') %>% 
  dplyr::mutate(Patient = factor(Patient))

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))

ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) + geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
  plot.theme
```

@@

dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
# Tiny heatmaps
plist = purrr::map(sum.patient.data$Patient, function(pt) {
  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

@@

@@

val.info

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
residuals = do.call(bind_rows, purrr::map(list.files(vdir,'residuals.tsv',recursive = T, full.names = T), function(f) read_tsv(f, col_types = 'ccddddl')))
failedQC = filter(residuals, varMAD_median > 0.008)
```

@@

all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
    return(pd)
})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')

@@

@@

```{r}
roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Validation cohort')
pROC::ci.auc(roc$auc)

ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)

@@

ggplot(melt(val.counts, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()
```

@@

val.preds = do.call(bind_rows, 
                    lapply(unique(val.preds$Patient) %>% 
                            map( function(p) val.preds %>% filter(Patient == p) %>% 
                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                 ), transform.by.time, info=val.info) )

@@

```{r, fig.width=6, fig.height=8}
require(survival)
require(survminer)
require(GGally)

@@

```{r}
vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% 
  mutate(Patient = as.factor(Patient)) %>% 
  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)

@@

merged %>%

@@

ggsave('plots/risk_km.png', plot=print(kmp), height=8, width=7, units = 'in', dpi = 300)

@@

max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {
    preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)
    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>% mutate(Endoscopy = Endoscopy*-1)
    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )

    left_join(rx, endo %>% dplyr::select(Endoscopy, months.before.final, Endoscopy.Year, Probability), by=c('Time 1'='Endoscopy')) %>%
      mutate(`Time 2` = ifelse(months.before.final <= 0,'Endpoint',`Time 2`)) %>% 
      mutate(Patient = pt) %>% mutate_at( vars(matches('Rule|Time')), funs(as.character) ) %>%
      dplyr::select(Patient, months.before.final, Endoscopy.Year, Probability, everything()) 
})
names(max.rules) = unique(pg.samp$Patient)

@@

pids = names(which(sapply(progressors, function(x) {
  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
})))

nids = names(which(sapply(nonprog, function(x) {
  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
})))

@@

```{r}
pids = names(which(sapply(progressors, function(x) {
  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
})))

nids = names(which(sapply(nonprog, function(x) {
  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
})))

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)
library(ggrepel)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(RColorBrewer)
library(ggfortify)

@@

cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()

# fit = glm(Status ~ age.diagnosis+M+gender, data=sum.patient.data, family = binomial(link='logit'))
# summary(fit)
# 
# sum.patient.data = sum.patient.data %>% mutate(total.followup = end.year - start.year)
# 
# surv =  Surv(sum.patient.data$total.followup,sum.patient.data$Status)
# terms = as.formula(paste("Surv(total.followup, status) ~", paste(c('age.diagnosis','M','gender'),collapse=" + "), sep=""))
# 
# cox = coxph(terms, data=sum.patient.data, na.action=na.omit,ties='breslow')
# summary(cox)
# 
# ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F, CI=T ) +
#       annotate("text", x=5, y=0.25,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),3),"\n", nrow(data), 'patients')) + theme_minimal()
# 
# 
# t.test(filter(sum.patient.data, Status == 'NP')$age.diagnosis,
#        filter(sum.patient.data, Status == 'P')$age.diagnosis)
# t.test(filter(sum.patient.data, Status == 'NP')$M, filter(sum.patient.data, Status == 'P')$M)
# fisher.test(table(sum.patient.data$Status, sum.patient.data$smoking))
# fisher.test(table(sum.patient.data$Status, sum.patient.data$gender))

@@

load(file, verbose=F)
melted = melt(dysplasia.df)

#melted$brks = as.integer(factor(cut(melted$value, breaks = quantile(melted$value, probs=c(seq(0,1,0.05))), include.lowest = T)))
p = ggplot(melted, aes(Var2, Var1, fill = value)) + geom_tile() +
  scale_fill_gradient2() + 
  theme(axis.text = element_blank()) + labs(x='Feature', y='sWGS Sample')
leg = get_legend(p)
ggsave('plots/samples.png', plot = p + theme(legend.position = 'none'), width=3, height=4, units='in')

@@

```{r echo=F, fig.height=5, fig.width=5}

autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

@@

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) #%>% arrange(-hazards)

p = ggplot(features, aes(label,hazards, fill=coef)) + geom_col() + 
  scale_fill_gradient2() + coord_flip() + theme_classic() + 
  theme(axis.ticks.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), legend.position = 'none') + labs(x='', y='', title='Effects') 
ggsave('plots/effects.png', plot=p, width = 3, height=1.5, units='in')

@@

error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv', col_types = 'cdD')

@@

#error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
#summary(error$Abs.Error)
#sd(error$Abs.Error)

p = ggplot(dplyr::select(pg.samp, Status, Samplename, RR), aes(reorder(Samplename, -RR), 1, fill=RR)) + geom_tile() + 
  #scale_fill_gradientn(colors=myPal) +
  scale_fill_gradient2(midpoint = 5, low = scales::muted('blue'), high = scales::muted('red')) + 
  scale_y_continuous(expand=c(0,0)) + 
  coord_flip() + labs(y='', x='') + theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.position = 'none') 
ggsave('plots/per_sample_RR.png', plot = p, width=0.5, height=3, units='in')

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC')

rocNOHGD$auc
pROC::ci.auc(rocNOHGD$auc)

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC'), width=6, height=6, units='in', dpi=300)
```

@@

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
#ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

@@

@@

# Not sure why I did this
# df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
# thresholds = 
#   rbind.data.frame(
#     est=t(round(pROC::coords(roc, "best", transpose = T),2)),
#     max=t(round(pROC::coords(roc, 0.8, transpose = T),2)),
#     'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose = T),2)),
#     'All samples'=t(round(pROC::coords(roc, 0.6, transpose = T),2))
# )
# thresholds$label=rownames(thresholds)
# 
# ggplot() + 
#   geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

@@

endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
  bind_rows(
    pROC::coords(rocEndoM, 'best', transpose = T),
    pROC::coords(rocEndoA, 'best', transpose = T),
    pROC::coords(rocLastEndo, 'best', transpose = T),
    pROC::coords(rocFirstEndo, 'best', transpose = T),
    pROC::coords(rocRandEndo, 'best', transpose = T)
  ))
```

@@

```{r}
pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time)) 
sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))

@@

rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

@@

@@

dir.create('plots/heatmaps/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/NP',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
# Tiny heatmaps
plist = purrr::map(sum.patient.data$Patient, function(pt) {
  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

@@

val.demo = readxl::read_xlsx(patient.file, 'Demographics') %>% 
  mutate_at(vars(Status,Sex), list(~factor(.))) %>%
  mutate(Followup = as.double(difftime(`Last Endoscopy`, `Initial Diagnosis`, units='weeks'))/52)

no_followup = filter(val.demo, Followup == 0 & Status == 'NP')
val.info = val.info %>% filter(Patient !=  no_followup$Patient)

val.demo = val.demo %>% mutate_at(vars(Sex,Smoking), list(~forcats::fct_explicit_na(.)))

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% tally %>% spread(Status, n)
val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Sex) %>% tally %>% spread(Status, n)

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Smoking) %>% tally %>% spread(Status, n)

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Maximal), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(`Age at diagnosis`), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Followup), list(~mean(.,na.rm=T), ~sd(.,na.rm=T))) 
val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Followup), list(~min(.,na.rm=T), ~max(.,na.rm=T)))

wilcox.test(filter(val.demo, Status == 'NP')$`Age at diagnosis`, filter(val.demo, Status == 'P')$`Age at diagnosis`)
wilcox.test(filter(val.demo, Status == 'NP')$Maximal, filter(val.demo, Status == 'P')$Maximal)

fisher.test(table(val.demo$Status, val.demo$Sex))
fisher.test(table(val.demo$Status, val.demo$Smoking))

# samples
val.info %>% group_by(Status) %>% tally

@@

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
residuals = do.call(bind_rows, purrr::map(list.files(vdir,'residuals.tsv',recursive = T, full.names = T), function(f) read_tsv(f, col_types = 'ccddddl')))
failedQC = filter(residuals, varMAD_median > 0.008)

@@

all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
    return(pd)
})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>%
  filter(`Hospital Research ID` != no_followup$`Hospital Research ID`)

@@

# samples passed QC
val.preds %>% group_by(Status) %>% tally

@@

```{r}
roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Validation cohort')
pROC::ci.auc(roc$auc)
ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)

@@

ggplot(melt(val.counts, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()

@@

val.preds = do.call(bind_rows, 
                    lapply(unique(val.preds$Patient) %>% map( function(p) val.preds %>% filter(Patient == p) %>% 
                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                 ), transform.by.time) )

@@

```{r, fig.width=6, fig.height=8, eval=F}
require(survival)
require(survminer)
require(GGally)

@@

```{r, eval=F}
vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% 
  mutate(Patient = as.factor(Patient)) %>% 
  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)

@@

#merged %>%

@@

kmp
ggsave('plots/risk_km.png', plot=print(kmp), height=8, width=7, units = 'in', dpi = 300)

@@

#pts = as.character(pg.samp %>% group_by(Patient, Endoscopy.Year) %>% tally %>% filter(n > 1) %>% dplyr::select(Patient) %>% distinct %>% pull)

max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {

    preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)

    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>%
      mutate(Endoscopy = as.integer(factor(Endoscopy*-1)))

    if (nrow(endo) == 1) return(NA)

    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )

    left_join(rx, endo %>% dplyr::select(Endoscopy, months.before.final, Endoscopy.Year, Probability), by=c('Time 1'='Endoscopy')) %>%
      mutate(`Time 2` = ifelse(months.before.final <= 0,'Endpoint',`Time 2`)) %>% 
      mutate(Patient = pt) %>% mutate_at( vars(matches('Rule|Time')), funs(as.character) ) %>%
      dplyr::select(Patient, months.before.final, Endoscopy.Year, Probability, everything()) 
})
names(max.rules) = unique(pg.samp$Patient)
max.rules = max.rules[which(sapply(max.rules, is.data.frame))]

@@

pids = names(which(unlist(  sapply(progressors, function(x) { nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2']) }) )))
nids = names(which(unlist(  sapply(nonprog, function(x) { nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2']) }) )))

@@

```{r}
pids = names(which(unlist(sapply(progressors, function(x) {
  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
}))))

nids = names(which(unlist(sapply(nonprog, function(x) {
  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
}))))

@@

p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`) %>% melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
  ggplot(aes(variable, value, group = Status)) + 
  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
p
ggsave('plots/second_endo_rx.png', plot=p, width=6, height=4, units = 'in', dpi = 300)
